% A LaTeX template for MSc Thesis submissions to 
% Politecnico di Milano (PoliMi) - School of Industrial and Information Engineering
%
% S. Bonetti, A. Gruttadauria, G. Mescolini, A. Zingaro
% e-mail: template-tesi-ingind@polimi.it
%
% Last Revision: October 2021
%
% Copyright 2021 Politecnico di Milano, Italy. NC-BY

\documentclass{Configuration_Files/PoliMi3i_thesis}

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------

% CONFIGURATIONS
\usepackage{parskip} % For paragraph layout
\usepackage{setspace} % For using single or double spacing
\usepackage{emptypage} % To insert empty pages
\usepackage{multicol} % To write in multiple columns (executive summary)
\setlength\columnsep{15pt} % Column separation in executive summary
\setlength\parindent{10pt} % Indentation
\raggedbottom  

% PACKAGES FOR TITLES
\usepackage{titlesec}
% \titlespacing{\section}{left spacing}{before spacing}{after spacing}
\titlespacing{\section}{0pt}{3.3ex}{2ex}
\titlespacing{\subsection}{0pt}{3.3ex}{1.65ex}
\titlespacing{\subsubsection}{0pt}{3.3ex}{1ex}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[english]{babel} % The document is in English  
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding
\usepackage[11pt]{moresize} % Big fonts

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\usepackage{svg}
\usepackage{transparent} % Enables transparent images
\usepackage{eso-pic} % For the background picture on the title page
%\usepackage{subfig} % Numbered and caption subfigures using \subfloat.
\usepackage[export]{adjustbox}
\usepackage{tikz} % A package for high-quality hand-made figures.
\usetikzlibrary{}
\graphicspath{{./Images/}} % Directory of the images
\usepackage{caption} % Coloured captions
\usepackage{subcaption}
\usepackage{xcolor} % Coloured captions
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{float}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{cancel}
\usepackage[overload]{empheq} % For braced-style systems of equations.
\usepackage{fix-cm} % To override original LaTeX restrictions on sizes

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % Tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
\bibliographystyle{abbrvnat} % You may use a different style adapted to your field

% OTHER PACKAGES
\usepackage{pdfpages} % To include a pdf file
\usepackage{afterpage}
\usepackage{lipsum} % DUMMY PACKAGE
\usepackage{fancyhdr} % For the headers
\usepackage{fancyvrb}
\usepackage[acronym]{glossaries}
\usepackage{enumitem} 
\fancyhf{}

\usepackage{listings}

\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

% Input of configuration file. Do not change config.tex file unless you really know what you are doing. 
\input{Configuration_Files/config}

%----------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%----------------------------------------------------------------------------

% EXAMPLES OF NEW COMMANDS
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation

%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------

%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------

\input{Thesis_Acronyms}

%----------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%----------------------------------------------------------------------------

\begin{document}

\fancypagestyle{plain}{%
\fancyhf{} % Clear all header and footer fields
\fancyhead[RO,RE]{\thepage} %RO=right odd, RE=right even
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%----------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\puttitle{
	title=Automated generation of efficient fuzzers for games using libFuzzer and RuleBook, % Title of the thesis
	name=Cem Cebeci, % Author Name and Surname
	course=Computer Science and Engineering \\Ingegneria Informatica, % Study Programme (in Italian)
	ID  =  10837160,  % Student ID number (numero di matricola)
	advisor= Prof. Giovanni Agosta, % Supervisor name
	coadvisor= Massimo Fioravanti, % Co-Supervisor name, remove this line if there is none
	academicyear={2023-2024},  % Academic Year
} % These info will be put into your Title page 

%----------------------------------------------------------------------------
%	PREAMBLE PAGES: ABSTRACT (inglese e italiano), EXECUTIVE SUMMARY
%----------------------------------------------------------------------------
\pagebreak

\startpreamble
\setcounter{page}{1} % Set page counter to 1

% ABSTRACT IN ENGLISH
\chapter*{Abstract} 
Fuzzing is the practice of repeatedly calling a program with semi-random inputs.
It is a popular technique to discover bugs and vulnerabilities in software.
However, fuzzing is not very practical when applied to input intensive programs such as games.
A major factor in this is that games often have complex input patterns.
They do not operate on a single large input but instead on a multitude of actions taken on them by the players.
These actions must be taken according to some potentially complex rules, a particular action is only available on some game states and only with certain parameters.
Therefore, even if a fuzzer knows all possible actions it can take on a game, generating valid sequences of actions can be a difficult task.
An invalid sequence of actions does not make an interesting fuzz input, since the game is expected to fail with such an input.
Consequently, frequently generating invalid sequences of actions decreases the fuzzer performance significantly.
Luckily, information about when an action can be taken and the constraints on its parameters should be available in the game logic, since the game needs to validate the actions players take and display error messages when invalid actions are taken.
Therefore, compiler techniques should be able to extract and exploit this information already present in the game rules to produce high-quality fuzzers.

In this work, we propose a method to generate fuzzers for games that utilize information about the dynamic availability of actions.
We describe a technique to parse an arbitrary sequence of bytes as a sequence of actions on the game, then detail how we minimize the number of invalid actions in the parsed action sequences.
We implement this technique in the RuleBook language compiler. RuleBook is a DSL for describing games and the RuleBook compiler has the information relevant to our technique available at compile time.
We measure the performance of four versions of fuzzers generated with this method, varying on how much information about the target game they utilize, along with two versions of fuzzers generated using OpenSpiel's game implementations.
Using three sample games, we demonstrate that the fuzzers generated from RL descriptions are as performant as their OpenSpiel counterparts.
Furthermore, we show that between RL fuzzers, versions that make the most use of the information RL exposes perform the best.
\\
\\
\textbf{Keywords:} compilers, fuzzing, games

% ABSTRACT IN ITALIAN
\chapter*{Abstract in lingua italiana}
Il fuzzing è la pratica di richiamare ripetutamente un programma con input semi-casuali.
È una tecnica popolare per scoprire bug e vulnerabilità nel software.
Tuttavia, il fuzzing non è molto pratico se applicato a programmi ad alta intensità di input come i giochi.
Un fattore importante è che i giochi hanno spesso modelli di input complessi.
Non operano su un singolo input di grandi dimensioni, ma su una moltitudine di azioni eseguite dai giocatori.
Queste azioni devono essere eseguite secondo regole potenzialmente complesse; una particolare azione è disponibile solo in alcuni stati del gioco e solo con determinati parametri.
Pertanto, anche se un fuzzer conosce tutte le possibili azioni che può compiere su un gioco, generare sequenze di azioni valide può essere un compito difficile.
Una sequenza di azioni non valida non costituisce un input interessante per il fuzz, poiché ci si aspetta che il gioco fallisca con tale input.
Di conseguenza, la generazione frequente di sequenze di azioni non valide riduce notevolmente le prestazioni del fuzzer.
Fortunatamente, le informazioni su quando un'azione può essere eseguita e i vincoli sui suoi parametri dovrebbero essere disponibili nella logica di gioco, poiché il gioco deve convalidare le azioni eseguite dai giocatori e visualizzare messaggi di errore quando vengono eseguite azioni non valide.
Pertanto, le tecniche di compilazione dovrebbero essere in grado di estrarre e sfruttare queste informazioni già presenti nelle regole del gioco per produrre fuzzer di alta qualità.

In questo lavoro proponiamo un metodo per generare fuzzer per giochi che utilizzano informazioni sulla disponibilità dinamica delle azioni.
Descriviamo una tecnica per analizzare una sequenza arbitraria di byte come una sequenza di azioni sul gioco, quindi spieghiamo come minimizzare il numero di azioni non valide nelle sequenze di azioni analizzate.
Questa tecnica è stata implementata nel compilatore del linguaggio RuleBook. RuleBook è un DSL per la descrizione dei giochi e il compilatore RuleBook ha a disposizione le informazioni rilevanti per la nostra tecnica al momento della compilazione.
Misuriamo le prestazioni di quattro versioni di fuzzer generati con questo metodo, che variano in base alla quantità di informazioni sul gioco di destinazione che utilizzano, insieme a due versioni di fuzzer generati utilizzando le implementazioni di gioco di OpenSpiel.
Utilizzando tre giochi campione, dimostriamo che i fuzzer generati dalle descrizioni RL hanno le stesse prestazioni delle loro controparti OpenSpiel.
Inoltre, dimostriamo che tra i fuzzer RL, le versioni che sfruttano maggiormente le informazioni esposte da RL sono le più performanti.
\\
\\
\textbf{Parole chiave:} compilers, fuzzing, giochi
%----------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES/SYMBOLS
%----------------------------------------------------------------------------

% TABLE OF CONTENTS
\thispagestyle{empty}
\tableofcontents % Table of contents 
\thispagestyle{empty}
\cleardoublepage

%-------------------------------------------------------------------------
%	THESIS MAIN TEXT
%-------------------------------------------------------------------------
% In the main text of your thesis you can write the chapters in two different ways:
%
%(1) As presented in this template you can write:
%    \chapter{Title of the chapter}
%    *body of the chapter*
%
%(2) You can write your chapter in a separated .tex file and then include it in the main file with the following command:
%    \chapter{Title of the chapter}
%    \input{chapter_file.tex}
%
% Especially for long thesis, we recommend you the second option.

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

% --------------------------------------------------------------------------
% NUMBERED CHAPTERS % Regular chapters following
% --------------------------------------------------------------------------
\chapter{Introduction}
\input{Introduction.tex}
\chapter{Background}
\input{Background.tex}
\chapter{State of the art}
\input{StateOfTheArt.tex}
\chapter{Solution design}\label{solutionDesign}
\input{SolutionDesign.tex}
\chapter{Evaluation}
\input{Evaluation.tex}
\chapter{Conclusion}
\input{Conclusion.tex}

%-------------------------------------------------------------------------
%	BIBLIOGRAPHY
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\bibliography{Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

%-------------------------------------------------------------------------
%	APPENDICES
%-------------------------------------------------------------------------

\cleardoublepage
\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
% \appendix
% \chapter{Appendix A}
% If you need to include an appendix to support the research in your thesis, you can place it at the end of the manuscript.
% An appendix contains supplementary material (figures, tables, data, codes, mathematical proofs, surveys, \dots)
% which supplement the main results contained in the previous chapters.

% \chapter{Appendix B}
% It may be necessary to include another appendix to better organize the presentation of supplementary material.

% LIST OF FIGURES
\listoffigures

% LIST OF TABLES
% \listoftables

% % LIST OF SYMBOLS
% % Write out the List of Symbols in this page
% \chapter*{List of Symbols} % You have to include a chapter for your list of symbols (
% \begin{table}[H]
%     \centering
%     \begin{tabular}{lll}
%         \textbf{Variable} & \textbf{Description} & \textbf{SI unit} \\\hline\\[-9px]
%         $\bm{u}$ & solid displacement & m \\[2px]
%         $\bm{u}_f$ & fluid displacement & m \\[2px]
%     \end{tabular}
% \end{table}

\cleardoublepage

\end{document}
